---
title: "Title of project"
author: "Anastasiia Revva, Juan Felipe Villamil Lozano, Aliona Hoste"
date: "17/11/2020"
output: html_document
---

### Introduction
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Objectives and hypotheses

### Methods 
What have you done? And what did you use to do it? Step by step, so anyone could do it again by reading this part.

We deleted accessions number 7 and 16 because last table 'Gas exchange parameters' does not contain informations


For calculating MFV, we used the same formula, however according to our method we used the min and max of the mean STI


### Results 
Figures and tables with captions and description of what do they mean.



```{r morphological traits significant difference}

boxplot(Shoot_Length ~ Treat_Contr, data = Morpho_t)
boxplot(Root_Length ~ Treat_Contr, data = Morpho_t)
boxplot(Height_Height ~ Treat_Contr, data = Morpho_t)
boxplot(Number_Leaves ~ Treat_Contr, data = Morpho_t)
boxplot(Leaf_Area ~ Treat_Contr, data = Morpho_t)


a1 <- aov(Shoot_Length ~ Treat_Contr, data = Morpho_t)
shapiro.test(a1$residuals)
#No normality
kruskal.test(Shoot_Length ~ Treat_Contr, data = Morpho_t)
# statistical difference

a2 <- aov(Root_Length ~ Treat_Contr, data = Morpho_t)
shapiro.test(a2$residuals)
#No normality
kruskal.test(Root_Length ~ Treat_Contr, data = Morpho_t)
# statistical difference


a3 <- aov(Plant_Height ~ Treat_Contr, data = Morpho_t)
shapiro.test(a3$residuals)
# No normality
kruskal.test(Plant_Height ~ Treat_Contr, data = Morpho_t)
# statistical difference


a4 <- aov(Number_Leaves ~ Treat_Contr, data = Morpho_t)
shapiro.test(a4$residuals)
# No normality 
kruskal.test(Number_Leaves ~ Treat_Contr, data = Morpho_t)
# statistical difference


a5 <- aov(Leaf_Area ~ Treat_Contr, data = Morpho_t)
shapiro.test(a5$residuals)
# no normality 
kruskal.test(Leaf_Area ~ Treat_Contr, data = Morpho_t)
# statistical difference

```





### Discussion 
From your objectives.

### Conclusions


