---
title: "Title of project"
author: "Anastasiia Revva, Juan Felipe Villamil Lozano, Aliona Hoste"
date: "17/11/2020"
output: html_document
---


### Introduction
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Objectives and hypotheses
Our main objectives are:

1. Explore the effects of salt stress on various morpho-physiological traits in different *E. sativa* accessions. 
2. Define the susceptible and resistant accessions
3. Identify traits the salt stress has the biggest influence on
4. Compare our results with the ones from the paper

Main hypothesis:

- Susceptible accessions have statistically different morpho-physiological traits compared to their control.
- The susceptible accessions have more morpho-physiological traits affected 


### Methods 
Firs we had to download and tidy the data.



####What have you done? And what did you use to do it? Step by step, so anyone could do it again by reading this part.

**
```{r Tidy_data_Nastya_experimenting.R, echo = TRUE, message=FALSE }
##Calling for the libraries to transform and tidy the data
library(dplyr)
library(tidyr)
##Downloading the data and transforming it 
#install.packages("ggplot2", dependencies = TRUE)
library(ggplot2)
#install.packages("readxl")
library(readxl)
# downloading excel sheet from internet and storing it into Excel object
Excel <- "https://dfzljdn9uc3pi.cloudfront.net/2020/9749/1/Raw_data_Afsar_et_al.%2C_2020-PeerJ_20.5.2020.xlsx"
# creating the folder datafile if it doesn't exist
D <- "data"
if (!dir.exists(D)) dir.create(D)
f <- paste0 (D, "/data.xlsx")
# downloading excel into the file datafile under the name data.xlsx
download.file(Excel, f, mode="wb")
# getting the name of excel sheets
excel.sheet <- excel_sheets(f)

# created a list with the 4 sheet table from excel 
X <- list()
for (i in excel.sheet){
  X[[i]] <- as.data.frame(read_xlsx(f,i))
}

# organise column names and fill the first 2 columns
for(i in names(X)){
  names(X[[i]])[1:3] <- c("Number", "Code", "Treat_Contr")
  X[[i]] <- X[[i]][-1,]
  #X[[i]] <- X[[i]][,-1] - deletion of numbers
  X[[i]] <- fill(X[[i]], 1:2)
}

#deleting the accession #7 and #16 from Morphological traits,
#Weight_ion, Chloro_c, because there is no data for that accessions in the Gas_e
#for(i in names(X)){
#  X[[i]] <- X[[i]][!X[[i]]$Code == "Es-7", ]
#  X[[i]] <- X[[i]][!X[[i]]$Code == "Es-16", ]
#}

#transforming columns data format 
for (i in seq_along(X)){
  X[[i]][2:3] <- lapply(X[[i]][2:3], as.factor)
  X[[i]][-(2:3)] <- lapply(X[[i]][-(2:3)], as.numeric)
}

# creating data frames from different xlsx sheets
Morpho_t <- X$`Morphological traits`
Weight_ion <- X$`FW DW RWC Ions EL`
Chloro_c <- X$`Chlorophyll content`
Gas_e <- X$`Gas Exchange parameters`

#Changing names of the columns for more convinient use
names(Morpho_t)[4:8] <- c('Shoot_Length', 
                          'Root_Length', 
                          'Plant_Height', 
                          'Number_Leaves', 
                          'Leaf_Area')
names(Chloro_c)[4] <- c('Chlorophyll_Content')
names(Weight_ion) [4:12] <- c('Fresh_Weight', 
                              'Dry_Weight', 
                              'RWC',
                              'Na',
                              'K', 
                              'Ca', 
                              'Mg', 
                              'K_Na', 
                              'Electrolyte_Leakage')
names(Gas_e) [4:7] <- c('Photsynthesis_Rate', 
                        'Intercellular_CO2', 
                        'Transpiration_Rate', 
                        'Stomatal_Conductance')

# adding NA cells into the missing 7th and 16th accession
# creating a data frame with the missing data first
Gas_e_miss <- data.frame (Number = rep(c(7, 16), each = 8),
                          Code = rep(c("Es-7", "Es-16"), each = 8),
                          Treat_Contr = rep(c("Treatment", "Control",
                                            "Treatment", "Control"), each = 4),
                          Photsynthesis_Rate = rep(NA, 16),
                          Intercellular_CO2 = rep(NA, 16),
                          Transpiration_Rate = rep(NA, 16),
                          Stomatal_Conductance = rep(NA, 16))
#changing the numbers in "Number" column
Gas_e$Number <- replace(Gas_e$Number, , rep(c(1:6, 8:15, 17:25), each = 8))
# adding NA cells into the missing 7th and 16th accession
if (!(is.element('Es-7', Gas_e$Code))) Gas_e <- rbind(Gas_e, Gas_e_miss)
#Sorting the variables according to Number
Gas_e <- Gas_e[order(Gas_e$Number),]
#Rewriting row names to correspond the ordered data
row.names(Gas_e) <- c(1:200)

# downloading data for the location of accessions
word <- "https://dfzljdn9uc3pi.cloudfront.net/2020/9749/1/Table_S1.docx"
w <- paste0(D, "/location.docx")
download.file(word, w, mode="wb")

#creating data frame for the location of accessions
#install.packages("docxtractr")
library(docxtractr)
Acc_loc <- docx_extract_all_tbls(read_docx(w, track_changes = NULL),
                                 guess_header = TRUE, preserve = FALSE,
                                 trim = TRUE)
Acc_loc <- as.data.frame(Acc_loc)
#removing m from Elevation column 
Acc_loc$Elevation <- sub(' m', '', Acc_loc$Elevation)
#transforming data in the accession_loc
Acc_loc[c(1,6)] <- lapply(Acc_loc[c(1,6)], as.numeric)
Acc_loc[-c(1,6)] <- lapply(Acc_loc[-c(1,6)], as.factor)
#In order to fit the length of this table to the length of the ultimate table
#we need to double all the rows
Acc_loc_double <- Acc_loc[rep(1:25, each = 2), ]

###
#experimenting with mapping the data
map <- data.frame(Code = Acc_loc$Code,
                  Latitude = Acc_loc$Latitude,
                  Longitude = Acc_loc$Longitude)
m <- ggplot(map, aes(x = Latitude,
                     y = Longitude))
m +
  geom_jitter() +
  geom_label(aes(label = Code))


# create summary table for morphological trait
a <- Morpho_t %>% 
  group_by(Number, Code, Treat_Contr) %>% 
  summarise(Shoot_Length = mean(Shoot_Length),
            Root_Length = mean(Root_Length), 
            Plant_Height = mean(Plant_Height), 
            Number_Leaves = mean(Number_Leaves), 
            Leaf_Area = mean(Leaf_Area))

# remove NA values from Weight_ion table
Weight_ion <- na.omit(Weight_ion)
#Create summary table for Weight ion
b <- Weight_ion %>% 
  group_by(Number, Code, Treat_Contr) %>% 
  summarise(Fresh_Weight= mean(Fresh_Weight),
            Dry_Weight= mean(Dry_Weight),
            RWC = mean(RWC),
            Na =mean(Na),
            K =mean(K),
            Ca =mean(Ca),
            Mg= mean(Mg),
            K_Na =mean(K_Na),
            Electrolyte_Leakage= mean(Electrolyte_Leakage))
#take only the columns number 4 to 12 to avoid repetition of number, code and treatment
b <- b[4:12] 

#Create summary table for Weight_ion
c <- Chloro_c %>% 
  group_by(Number, Code, Treat_Contr) %>% 
  summarise(Chlorophyll_Content = mean(Chlorophyll_Content))
#take only the columns number 4 to avoid repetition of number, code and treatment
c <- c[4]  

#Create summary table for Gas_e
d <- Gas_e %>% 
  group_by(Number, Code, Treat_Contr) %>% 
  summarise(Photsynthesis_Rate= mean(Photsynthesis_Rate),
            Intercellular_CO2= mean(Intercellular_CO2), 
            Transpiration_Rate= mean(Transpiration_Rate), 
            Stomatal_Conductance=mean(Stomatal_Conductance))
#take only the columns number 4 to 7 to avoid repetition of number, code and treatment
d <- d[4:7]

#add Acc_loc data to the table
e <- Acc_loc_double
e <- e[6]

#create data frame with all the summarized tables 
table <- data.frame(a, b, c, d, e)
```








We deleted accessions number 7 and 16 because last table 'Gas exchange parameters' does not contain informations


For calculating MFV, we used the same formula, however according to our method we used the min and max of the mean STI


### Results 
Figures and tables with captions and description of what do they mean.



```{r morphological traits significant difference}
a1 <- aov(Shoot_Length ~ Treat_Contr, data = Morpho_t)
shapiro.test(a1$residuals)
#No normality

a2 <- aov(Root_Length ~ Treat_Contr, data = Morpho_t)
shapiro.test(a2$residuals)
#No normality

a3 <- aov(Plant_Height ~ Treat_Contr, data = Morpho_t)
shapiro.test(a3$residuals)
# No normality

a4 <- aov(Number_Leaves ~ Treat_Contr, data = Morpho_t)
shapiro.test(a4$residuals)
# No normality 

a5 <- aov(Leaf_Area ~ Treat_Contr, data = Morpho_t)
shapiro.test(a5$residuals)
# no normality 

for (i in names(Morpho_t[4:8])) { 
  boxplot(Morpho_t[, i] ~ Morpho_t$Treat_Contr, 
          ylab = names(Morpho_t[i]), 
          xlab = "Treatment"
  )
  print(kruskal.test(Morpho_t[, i] ~ Morpho_t$Treat_Contr))
}


```

```{r mean value for salt-stressed }
#mean(x$Shoot_Length)
#mean(y$Shoot_Length)
#mean(x$Root_Length)
#mean(y$Root_Length)
#mean(x$Plant_Height)
#mean(y$Plant_Height)
#mean(x$Number_Leaves)
#mean(y$Number_Leaves)
#mean(x$Leaf_Area)
#mean(y$Leaf_Area)

#sd(x$Shoot_Length)
#sd(y$Shoot_Length)
#sd(x$Root_Length)
#sd(y$Root_Length)
#sd(x$Plant_Height)
#sd(y$Plant_Height)
#sd(x$Number_Leaves)
#sd(y$Number_Leaves)
#sd(x$Leaf_Area)
#sd(y$Leaf_Area)

#add JuanFe code from Analysis 1

```

There is a statistical difference between control and treatment for all morphological trait


- **SL**: Shoot length is significantly reduced in the salt-stressed plants compared with the control (p < 0.05). The mean value for SL in salt-stressed plants is 1.10 ± 0.48 and 2.30 ± 0.56 for control plants. SL stress index of accessions varied from 16.73 to 89.12

- **RL**: Root length is significantly reduced in the salt-stressed plants compared with the control (p < 0.05). The mean value for RL in salt-stressed plants is 0.98 ± 0.43 and 2.17 ± 0.42 for control plants. 
RL stress index of accessions varied from 15.91 to 73.53 

- **PH**: Plant height is significantly reduced in the salt-stressed plants compared with the control (p < 0.05).
The mean value for PH in salt-stressed plants is 2.18 ± 0.76 and 4.43 ± 0.77 for control plants. 
PH stress index of accessions varied from 19.83 to 74.07 

- **NL**: Number of leaves is significantly reduced in the salt-stressed plants compared with the control (p < 0.05).
The mean value for NL in salt-stressed plants is 2.43 ± 0.82 and 4.22 ± 0.78 for control plants. 
NL stress index of accessions varied from 28.00 to 94.12

- **LA**: Leaf area is significantly reduced in the salt-stressed plants compared with the control (p < 0.05).
The mean value for LA in salt-stressed plants is 0.03 ± 0.02 and 0.07 ± 0.03 for control plants. 
LA stress index of accessions varied from 75.95 to 100

### Discussion 
####From your objectives.










### Conclusions


